Measuring contrast processing in the visual system using SSVEP

### 2024-11-28 review.qmd
On 'binocular visual development'

The approach of estimating visual thresholds by extrapolating SSVEP responses at higher stimulus levels has revealed much about the development of visual abilities in infants [@Harris1976; @Braddick1986; @Atkinson1979]. In general, SSVEP measurements of infant vision have revealed that contrast sensitivity for both achromatic and chromatic contrast as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b] with both chromatic and achromatic contrast detection reaching near-adult levels by around six months. Spatial acuity as measured by SSVEP reaches adult levels more slowly, but near-adult levels are recorded around one year [@Norcia1985b; @Hamer1989] compared to around six to seven years with behavioural measures [@Ellemberg1999; @Atkinson1983]. At least some of this difference is likely due to the relative objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking, which require careful measurement of the infant's gaze direction. However it should be noted that other groups have reported electrophysiological correlates of visual acuity that more closely match the behavioural measures [@De-Vries-Khoe1982].

>>>>

The approach of estimating visual thresholds by extrapolating SSVEP responses at higher stimulus levels has revealed much about the development of visual abilities in infants [@Harris1976; @Braddick1986; @Atkinson1979]. In general, SSVEP measurements of infant vision have revealed that contrast sensitivity and acuity for both achromatic and chromatic stimuli as well as stereoscopic depth perception develops earlier than had been supposed previously based on behavioural readouts [@Dobson1978; @Norcia1985b]. For example, using SSVEPs, Norcia and Tyler [@Norcia1985b] found acuity thresholds in 1 month old infants that were roughly half those estimated previously from behavioural experiments (reviewed in [@brownDevelopmentVisualSensitivity1990a]).
SSVEP measures of thresholds for detecting purely chromatic stimuli find responses to isoluminant stimuli by around 5 to 6 weeks with contrast detection reaching near-adult levels by around six months [@morroneDevelopmentInfantContrast1993]. Spatial acuity as measured by SSVEP reaches adult levels more slowly, but near-adult levels are recorded around one year [@Norcia1985b; @Hamer1989] compared to around six to seven years with behavioural measures [@Ellemberg1999; @Atkinson1983]. At least some of this difference is likely due to the relative objectivity and high SNR of the SSVEP technique compared to other methods such as preferential looking, which require careful measurement of the infant's gaze direction.  In slightly older chilren, other groups have reported electrophysiological correlates of visual acuity that more closely match the behavioural measures [@De-Vries-Khoe1982].

The ability to use information from both eyes
